### 2025-01-10 documentation/supplement.tex
final updates before resubmission

\subsection{Safety}
Class 4 lasers are extremely dangerous. A reflection on a surface can be enough to inflict permanent eye damage. Safety protocols must be observed. Consult with your institution's environmental health and safety officer and take proper training before performing any work with a laser system. Learn from people experienced with these devices and work carefully.\newline
During alignment, always wear goggles (ours were ML1 by NoIR LaserShields with an optical density $>7$ in the relevant range around $920\text{ nm}$), remove jewelry and watches, and use the low laser power (“alignment mode”) whenever possible. Once the assembly progresses further, full power mode is necessary. Extra care must be taken then. Always use Laser Viewing Cards to track the laser. We use Thorlabs VRC4. Always be conscious where the beam is, and where the beam goes. The beam must be terminated in beam blockers and should be shielded whenever possible with opaque plastic or anodized aluminium tubes to increase safety.

@@

\subsection{Structure and use of this document}
In the following sections, we will cover the setup and alignment of the microscope. This is split into three parts. The first part covers all components on the optical table that serve to condition the beam. The second part focuses on the  95 mm vertical post that holds the scanning mirror and scan lens assemblies. The third part is the microscope head with the attached collection optics. We will begin with the assembly on the optical table. All along, we assume the reader will have downloaded and opened the CAD model as we will rely heavily on it. Further, check carefully the instruction manuals of the scanning mirrors, photomultiplier tubes (PMTs), power supplies, Pockels cell, ScanImage, the Piezoelectric objective collar, and the filter mounting instructions.\newline
Before building your own microscope, read these instructions fully. Check the CAD files, and make sure that the instructions make sense. This document follows closely our own alignment procedure, and we have done our best to be accurate. However, use common sense. Do not follow these instructions blindly (or you might follow these instructions blindly). Like all open instrumentation (under CC-BY-4.0), this work is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY, to the extent permitted by law; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.

@@

  \caption{{\bf Scanner electronics.} \textbf{A)} Overview of all scanning mirror parts. Key parts are labeled with capital letters. Optomechanics were covered in Fig.~\ref{sfig10}. All electrical drivers were placed in the yellow circuit box. \textbf{B)} Power supply and connection of the scanning mirrors with key parts labeled by capital letters. \textcolor{blue}{\textbf{C)} Wiring diagram for the $\pm 28\text{ V}$ power supply, using two galvanically isolated supplies. \textbf{D)} The scanning mirror driver boards, provided by Novanta, and mounted in the yellow circuit box. The smaller driver circuit (blue) powers the resonant scanner. The larger circuit board (green) drives the slow galvanometer. The circuit boards are mounted to the aluminium walls with screws and heat transfer compound. \textbf{E)} Same box as in D, but from the outside and key parts labeled by capital letters.}}

@@

\textcolor{blue}{To provide these three signals, we used a power supply with two galvanically isolated and adjustable output channels, that can be connected in series (Note: not every power supply supports this function. We used a GPD-3303D from Gw Instek that allows this function with the SER/INDEP key). The voltage of each of the two unconnected channels was set to +28 V. To produce the three supply signals, we used these two channels in series. In other words, we connect the positive terminal of the first channel to the negative terminal of the second channel. As the channels are otherwise not connected to a common ground (this is meant with the words ``galvanically isolated''), this does not produce a short circuit. Instead, this link effectively provides a new reference point, relative to which the first supply produces a negative -28 V supply, and the second supply produces the +28 V supply. This is illustrated in Fig.~\ref{sfig11}C. In this arrangement, a multimeter would read between points G and I +28 V, between G and F -28 V, and between F and I +56 V. The supply of the resonant scanners is simpler, as it only requires a single +12 V line. The supplies are connected to the power supply ports on the prototype box, in which the two driver boards reside, see Fig.~\ref{sfig11}D. Signal lines are connected on the outside with coaxial cables to the MBF Bioscience DAQ system, see Fig.~\ref{sfig11}E.} Key components in the panels are labeled with capital letters. These are as follows:

@@

\section{Recommendations for use}
Make a point of regularly checking your microscope. Lenses can get dirty, the path misaligned, and components age.
%
\begin{itemize}
    \item Measure (1) the laser's power levels, (2) the Pockels cell calibration and (3) check exposed lenses and mirrors for dust and fingerprints with a flashlight. This is fast and should be done before every experiment.
    \item When the microscope is not in use, it is good practice to switch off all components, except the scanning mirror power supplies, the PMT preamplifiers, and the photodiode.
    \item Every once in a while, switch the red and green PMTs to check for signs of aging and validate their performance on a fluorescent slide or sample of known brightness. The PMTs can degrade and may need to be replaced, depending on use, after several years. They should be treated with care: power ramped up and down slowly, and saturation should be avoided. We use a rotation system in which new PMTs are typically used for the green channel. When their performance becomes worse, they often have a few more months or years in the (typically brighter) red channel in them.
    \item It is recommended to periodically image a bead sample to validate the PSF, as well as a Fluorescein bath to detect possible misalignment. For a detailed protocol, see Lees et al. \cite{Lees2024}.
    \item Be mindful of imaging artifacts. When suspecting an issue, go back to the basics, and isolate the error.
\end{itemize}
%
 For safety, it is good practice to shield the laser path. In addition to safety concerns, it will also help in keeping dust and dirt away from optical components.

\section{Deriving the relationship between intensity and dispersion}
\textcolor{blue}{In the main article, we stated that the expected fluorescence $I(\varphi)$ as function of group delay dispersion $\varphi$ applied to a $\varphi_0$-dispersed Gaussian pulse of full-width-at-half-maximum $\Delta\tau$ can be written as
\begin{eqnarray}
    I(\varphi)=I_0 \left(1+16\log(2)^2\frac{(\varphi+\varphi_0)^2}{\Delta\tau^4}\right)^{-1/2}\label{main}.
\end{eqnarray}
To derive this expression, consider the electrical field of a Gaussian pulse of amplitude $E_0$, width $\Delta t$, center frequency $\omega_0$, phase $\Theta$, and centered at time $t=0$:
\begin{eqnarray}
E(t)=E_0\exp\left(-\frac{t^2}{4\Delta t^2}\right)\cdot\exp\left(-\complexi\left(\omega_0 t + \Theta\right)\right).
\end{eqnarray}
Its intensity is 
\begin{eqnarray}
P(t) = |E(t)|^2 = |E_0|^2 \exp\left(-\frac{t^2}{2\Delta t^2}\right)\label{intens}.
\end{eqnarray}
In the Fourier domain, the spectral content of this wave package is:
\begin{eqnarray}
\tilde{E}(\omega)=E_0\cdot\sqrt{2}\Delta t\, \exp\left(-\complexi\,\Theta -\Delta t^2(\omega - \omega_0)^2\right).
\end{eqnarray}
In the linear regime, frequency-dependent dispersion changes the electric field to
\begin{eqnarray}
\tilde{E}(\omega)\to\tilde{E}(\omega)\cdot\exp\left(\complexi\,\phi(\omega-\omega_0)\right)
\end{eqnarray}
with phase shift $\phi(\omega-\omega_0)$. We approximate this function to quadratic order $\phi(\omega-\omega_0)\approx a + b(\omega-\omega_0) + \frac{c}{2}(\omega-\omega_0)^2$. With this approximation, we obtain the Fourier representation of the electric field of the dispersed pulse:
\begin{eqnarray}
\tilde{E}^\text{disp}(\omega)=\tilde{E}(\omega)\cdot\exp\left(\complexi\,\left(a + b(\omega-\omega_0) + \frac{c}{2}(\omega-\omega_0)^2\right)\right).
\end{eqnarray}
Coefficient $a$ causes a phase shift and coefficient $b$ a time shift. Both do not change pulse width. Computing the inverse Fourier transform to obtain $E^\text{disp}(t)$ allows to compute the intensity of the dispersed pulse as 
\begin{eqnarray}
P^\text{disp}(t) = |E^\text{disp}(t)|^2 = |E_0|^2\frac{2\Delta t^2}{\sqrt{c^2 + 4\Delta t^4}}\exp\left(-\frac{2(b+t)^2\Delta t^2}{c^2+4\Delta t^4}\right)\label{intens_disp}.
\end{eqnarray}
If the average power of the laser across dispersion compensation settings $c$ is constant, the prefactor is approximately constant. Comparison with Eq. \eqref{intens} then shows that the pulse has broadened in time to $\Delta t\to \frac{\sqrt{c^2+4\Delta t^4}}{2\Delta t}$. Rewriting the widths of the dispersed pulse in the more common full-width-at-half-maximum $\Delta\tau$, using $\Delta t \to \Delta\tau/\sqrt{8\log(2)}$ for a Gaussian pulse, we obtain
\begin{eqnarray}
\Delta\tau^\text{disp} = \frac{\sqrt{\Delta\tau^4 + 16\log(2)^2c^2}}{2 \Delta\tau\sqrt{\log(4)}}.
\end{eqnarray}
The two-photon fluorescence intensity $I$ is proportional to the inverse of the pulse width, and the group delay dispersion is composed of the original dispersion $\varphi_0$ and the applied correction $\varphi$, so that $c \to \varphi_0 + \varphi$. We obtain
\begin{eqnarray}
I(\varphi) \propto \frac{1}{\Delta\tau^\text{disp}} = \frac{2 \Delta\tau\sqrt{\log(4)}}{\sqrt{\Delta\tau^4 + 16\log(2)^2(\varphi_0 + \varphi)^2}} = \frac{2\sqrt{\log(4)}}{\Delta\tau}\frac{1}{\sqrt{1 + 16\log(2)^2\frac{(\varphi_0 + \varphi)^2}{\Delta\tau^4}}}.
\end{eqnarray}
Absorbing the constants $\frac{2\sqrt{\log(4)}}{\Delta\tau}\to I_0$ yields Eq.~\eqref{main}. The same result for pulse broadening, with $16\log(2)^2\approx 7.68\dots$, is stated as Eq.~(3) in \cite{Zipfel2003}.
}
\nolinenumbers

>>>>

\subsection{Safety}
Class 4 lasers are extremely dangerous. A reflection on a surface can be enough to inflict permanent eye damage. Safety protocols must be observed. Consult with your institution's environmental health and safety officer and take proper training before performing any work with a laser system. Learn from people experienced with these devices and work carefully.\newline
During alignment, always wear goggles (ours were ML1 by NoIR LaserShields with an optical density $>7$ in the relevant range around $920\text{ nm}$), remove jewelry and watches, only use-reflective tools, do not store unnecessary items on the optical table, and use the low laser power (“alignment mode”) whenever possible. Once the assembly progresses further, full power mode is necessary. Extra care must be taken then. Always use Laser Viewing Cards to track the laser. We use Thorlabs VRC4. Always be conscious where the beam is, and where the beam goes. The beam must be terminated in beam blockers and should be shielded whenever possible with opaque plastic or anodized aluminium tubes to increase safety.

@@

\subsection{Structure and use of this document}
In the following sections, we will cover the setup and alignment of the microscope. This is split into three parts. The first part covers all components on the optical table that serve to condition the beam. The second part focuses on the  95 mm vertical post that holds the scanning mirror and scan lens assemblies. The third part is the microscope head with the attached collection optics. We will begin with the assembly on the optical table. All along, we assume the reader will have downloaded and opened the CAD model as we will rely heavily on it. Further, check carefully the instruction manuals of the scanning mirrors, photomultiplier tubes (PMTs), power supplies, Pockels cell, ScanImage, the Piezoelectric objective collar, and the filter mounting instructions.\newline
Before building your own microscope, read these instructions fully. Check the CAD files, and make sure that the instructions make sense. This document follows closely our own alignment procedure, and we have done our best to be accurate. However, use common sense. Do not follow these instructions blindly (or you might follow these instructions blindly in the future). Like all open instrumentation (under CC-BY-4.0), this work is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY, to the extent permitted by law; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.

@@

  \caption{{\bf Scanner electronics.} \textbf{A)} Overview of all scanning mirror parts. Key parts are labeled with capital letters. Optomechanics were covered in Fig.~\ref{sfig10}. All electrical drivers were placed in the yellow circuit box. \textbf{B)} Power supply and connection of the scanning mirrors with key parts labeled by capital letters. \textcolor{black}{\textbf{C)} Wiring diagram for the $\pm 28\text{ V}$ power supply, using two galvanically isolated supplies. \textbf{D)} The scanning mirror driver boards, provided by Novanta, and mounted in the yellow circuit box. The smaller driver circuit (blue) powers the resonant scanner. The larger circuit board (green) drives the slow galvanometer. The circuit boards are mounted to the aluminium walls with screws and heat transfer compound. \textbf{E)} Same box as in D, but from the outside and key parts labeled by capital letters.}}

@@

\textcolor{black}{To provide these three signals, we used a power supply with two galvanically isolated and adjustable output channels, that can be connected in series (Note: not every power supply supports this function. We used a GPD-3303D from Gw Instek that allows this function with the SER/INDEP key). The voltage of each of the two unconnected channels was set to +28 V. To produce the three supply signals, we used these two channels in series. In other words, we connect the positive terminal of the first channel to the negative terminal of the second channel. As the channels are otherwise not connected to a common ground (this is meant with the words ``galvanically isolated''), this does not produce a short circuit. Instead, this link effectively provides a new reference point, relative to which the first supply produces a negative -28 V supply, and the second supply produces the +28 V supply. This is illustrated in Fig.~\ref{sfig11}C. In this arrangement, a multimeter would read between points G and I +28 V, between G and F -28 V, and between F and I +56 V. The supply of the resonant scanners is simpler, as it only requires a single +12 V line. The supplies are connected to the power supply ports on the prototype box, in which the two driver boards reside, see Fig.~\ref{sfig11}D. Signal lines are connected on the outside with coaxial cables to the MBF Bioscience DAQ system, see Fig.~\ref{sfig11}E.} Key components in the panels are labeled with capital letters. These are as follows:

@@

\section{Recommendations for use}
Make a point of regularly checking your microscope. Lenses can get dirty, the path misaligned, and components age.
%
\begin{itemize}
    \item Measure (1) the laser's power levels, (2) the Pockels cell calibration and (3) check exposed lenses and mirrors for dust and fingerprints with a flashlight. This is fast and should be done before every experiment.
    \item When the microscope is not in use, it is good practice to switch off all components, except the scanning mirror power supplies, the PMT preamplifiers, and the photodiode.
    \item Every once in a while, switch the red and green PMTs to check for signs of aging and validate their performance on a fluorescent slide or sample of known brightness. The PMTs can degrade and may need to be replaced, depending on use, after several years. They should be treated with care: power ramped up and down slowly, and saturation should be avoided. We use a rotation system in which new PMTs are typically used for the green channel. When their performance becomes worse, they often have a few more months or years in the (typically brighter) red channel in them.
    \item It is recommended to periodically image a bead sample to validate the PSF, as well as a Fluorescein bath to detect possible misalignment. For a detailed protocol, see Lees et al. \cite{Lees2024}.
    \item Be mindful of imaging artifacts. When suspecting an issue, go back to the basics, and isolate the error.
\end{itemize}
%
 For safety, it is good practice to shield the laser path. In addition to safety concerns, it will also help in keeping dust and dirt away from optical components.
\newpage

\section{Deriving the relationship between intensity and dispersion}
\textcolor{black}{We stated that the expected fluorescence $I(\varphi)$ as function of group delay dispersion $\varphi$ applied to a $\varphi_0$-dispersed Gaussian pulse of full-width-at-half-maximum $\Delta\tau$ can be written as
\begin{eqnarray}
    I(\varphi)=I_0 \left(1+16\log(2)^2\frac{(\varphi+\varphi_0)^2}{\Delta\tau^4}\right)^{-1/2}\label{main}.
\end{eqnarray}
To derive this expression, consider the electrical field of a Gaussian pulse of amplitude $E_0$, width $\Delta t$, center frequency $\omega_0$, phase $\Theta$, and centered at time $t=0$:
\begin{eqnarray}
E(t)=E_0\exp\left(-\frac{t^2}{4\Delta t^2}\right)\cdot\exp\left(-\complexi\left(\omega_0 t + \Theta\right)\right).
\end{eqnarray}
Its intensity is 
\begin{eqnarray}
P(t) = |E(t)|^2 = |E_0|^2 \exp\left(-\frac{t^2}{2\Delta t^2}\right)\label{intens}.
\end{eqnarray}
In the Fourier domain, the spectral content of this wave package is:
\begin{eqnarray}
\tilde{E}(\omega)=E_0\cdot\sqrt{2}\Delta t\, \exp\left(-\complexi\,\Theta -\Delta t^2(\omega - \omega_0)^2\right).
\end{eqnarray}
In the linear regime, frequency-dependent dispersion changes the electric field to
\begin{eqnarray}
\tilde{E}(\omega)\to\tilde{E}(\omega)\cdot\exp\left(\complexi\,\phi(\omega-\omega_0)\right)
\end{eqnarray}
with phase shift $\phi(\omega-\omega_0)$. We approximate this function to quadratic order $\phi(\omega-\omega_0)\approx a + b(\omega-\omega_0) + \frac{c}{2}(\omega-\omega_0)^2$. With this approximation, we obtain the Fourier representation of the electric field of the dispersed pulse:
\begin{eqnarray}
\tilde{E}^\text{disp}(\omega)=\tilde{E}(\omega)\cdot\exp\left(\complexi\,\left(a + b(\omega-\omega_0) + \frac{c}{2}(\omega-\omega_0)^2\right)\right).
\end{eqnarray}
Coefficient $a$ causes a phase shift and coefficient $b$ a time shift. Both do not change pulse width. Computing the inverse Fourier transform to obtain $E^\text{disp}(t)$ allows to compute the intensity of the dispersed pulse as 
\begin{eqnarray}
P^\text{disp}(t) = |E^\text{disp}(t)|^2 = |E_0|^2\frac{2\Delta t^2}{\sqrt{c^2 + 4\Delta t^4}}\exp\left(-\frac{2(b+t)^2\Delta t^2}{c^2+4\Delta t^4}\right)\label{intens_disp}.
\end{eqnarray}
Comparison with Eq. \eqref{intens} then shows that the pulse has broadened in time to $\Delta t\to \frac{\sqrt{c^2+4\Delta t^4}}{2\Delta t}$. Rewriting the widths of the dispersed pulse in the more common full-width-at-half-maximum $\Delta\tau$, using $\Delta t \to \Delta\tau/\sqrt{8\log(2)}$ for a Gaussian pulse, we obtain
\begin{eqnarray}
\Delta\tau^\text{disp} = \frac{\sqrt{\Delta\tau^4 + 16\log(2)^2c^2}}{2 \Delta\tau\sqrt{\log(4)}}.
\end{eqnarray}
The two-photon fluorescence intensity $I$ is proportional to the inverse of the pulse width, and the group delay dispersion is composed of the original dispersion $\varphi_0$ and the applied correction $\varphi$, so that $c \to \varphi_0 + \varphi$. We obtain
\begin{eqnarray}
I(\varphi) \propto \frac{1}{\Delta\tau^\text{disp}} = \frac{2 \Delta\tau\sqrt{\log(4)}}{\sqrt{\Delta\tau^4 + 16\log(2)^2(\varphi_0 + \varphi)^2}} = \frac{2\sqrt{\log(4)}}{\Delta\tau}\frac{1}{\sqrt{1 + 16\log(2)^2\frac{(\varphi_0 + \varphi)^2}{\Delta\tau^4}}}.
\end{eqnarray}
Absorbing the constants $\frac{2\sqrt{\log(4)}}{\Delta\tau}\to I_0$ yields Eq.~\eqref{main}. The same result for pulse broadening, with $16\log(2)^2\approx 7.68\dots$, is stated as Eq.~(3) in \cite{Zipfel2003}.
}

\nolinenumbers
